How to Schedule Periodic Messages in a Network Without Congestion

### 2017-06-01 article/article.tex
reprise du début et simplifications

%opening
\title{Contention Management for 5G}
\author{DB,CC,MG,OM,YS}

@@

\begin{abstract}
We study the sending of periodic messages in a graph following predefined paths, a problem arising from the communication between radio antennas and their signal processing units located in distant data-centers. The messages must be sent deterministically to minimize the latency. We prove that the problem is $\NP$-complete even for very restricted graphs. We then deal with the case of a star graph and for different regimes of the parameters we propose algorithms that are experimentally shown to work extremely well
compared to traditional statistical multiplexing.
\end{abstract}

@@

 The network topology is modeled by a directed weighted graph and a set of routes (paths) from source nodes (RRH) to target nodes (BBUs). The time is discretized and a unit of time corresponds to the transmission of a minimal unit of data over the network.
 The process must be periodic, that is if we look at the network at time $t$ and $t+P$ where $P$ is the period, all messages are at the same position in the networks. Moreover, we want to avoid any  buffering in internal nodes of the graph, therefore the periodic process we need to design cannot collision between packets. To ensure that we need to propose a deterministic process. 
 We have two sets of parameters that we can choose when building a periodic sending process or a \textbf{periodic assignment}, first the time at which each packet is sent by a RRH in each period and the waiting time in the BBU before the answer is sent back to the RRH. We have two major difficulties to overcome when building an periodic assignment. First the periodicity which makes many scheduling methods inapplicable, since one must ensure that a message does not collide with the other messages sent by the others BBU/RRH in the same period, but also in the previous and following periods. Second, one must minimize the latency, that is the time 
 between the emission of a message and the complete return of its answer. It means that the only buffering we are allowed in the BBU
 (the waiting time before sending back the answer) must be small, in particular when the route is long.

@@

\subsection*{Article organization}
 In Sec.~\ref{sec:def} we define our model, we formalize our problems PRA (one way periodic assignment) and PALL (two ways periodic assignment) and study their basic properties.
In Sec.~\ref{sec:complexity}, we prove that the problem PRA is $\NP$-hard and cannot be approximated even for very
restricted graphs. In the next two sections, we study the problem PALL case on a star topology which often arises in practice.
Since the problem is already quite complicated, we first study a variant called PAZL in Sec.\ref{sec:PAZL}, where the waiting times
are all zero. We provide several algorithms, some of their theoretical properties and give experimental evidences that they work well
when the network is no too loaded. Finally in Sec.\ref{sec:PALL}, we propose algorithms for the general problem and exhibit one which works extremely well in our experiments even in loaded networks. In particular we show that the deterministic communication schemes we design vastly outperform the traditional stochastic multiplexing.

\section{Model}\label{sec:def}

  \subsection{Definitions}

\todo{simplifier car les chemins sont toujours donnés dans l'article -> du coup le problème c'est un graphe + l'ensemble de routes
+ remplacer partout affectation par assignment -> simplifier au max et uniformiser les notations}

The network is modeled as a directed graph $G=(V,A)$. Each arc  $(u,v)$ in $A$ is labeled by an integer $dl(u,v) \geq 1$ that we call the delay and which represents the time slots taken by a signal to go from $u$ to $v$ using this arc. A {\bf route} $r$ in $G$ is a path, that is sequence of consecutive arcs $a_0, \ldots , a_{k-1}$, with $a_i=(u_i,u_{i+1}) \in A$.  The {\bf latency} of a vertex $u_i$ in a path $r$, with $i \geq 1$, is defined by $\lambda(u_i,r)= \sum\limits_{0 \leq j <i} dl(a_j)$. We also define $\lambda(u_0,r)=0$. The latency of the route $r$ is defined by $\lambda (r)= \lambda (u_k,r)$.

      A {\bf routing function} $\cal R$ in $G$ associates to each pair of vertices $(u,v)$ a route from $u$ to $v$. Let $\cal C$ be an {\bf assignment} in $G$, i.e., a set of couples of different vertices of $G$. We denote by $\cal R_{\cal C}$ the set of routes ${\cal R}(u,v)$ for any $(u,v)$ in $\cal C$. We call $\cal R_{\cal C}$ a {\bf routage graph}, it contains all the informations needed in the forthcoming problems (assignment, routes and delays of the arcs).

      \todo{Si on s'en sert, ajouter ici que le routage est cohérent.}

   \subsection{Slotted time Model}
      Consider now a positive integer $P$ called the {\bf period}. In our problem, we send messages in the network with period $P$. The time will thus be cut into slices of $P$ discrete slots. Assume we send a message at the source $s_i$ of the route $r_{s_i}$, at the time slot $m_{s_i}$ in the first period, then a message will be sent at time slot $m_{s_i}$ at each new period. We define the first time slot at which the message reaches a vertex $v$ in this route by $t(v,r_{s_i}) = m_{s_i} + \lambda(v,r) \mod P$.

      A message usually cannot be transported in a single time slot. We denote by $\tau$ the number 
      of consecutive slots necessary to transmit a message. Let us call $[t(v,r_{s_i})]_{P,\tau}$ the set of time slots used by $r_{s_i}$ at a vertex $v$ in a period $P$, that is $[t(v,r_{s_i})]_{P,\tau} = \{t(v,r_{s_i}) + k \mod P \mid 0 \leq k < \tau \}$. Usually $P$ and $\tau$ will be clear from the context and we will denote $[t(v,r_{s_i})]_{P,\tau}$ by $[t(v,r_{s_i})]$.

      A {\bf $(P,\tau)$-periodic affectation} of a routage graph $\cal R_{\cal C}$ is a sequence  ${\cal M}=(m_{s_0}, \ldots ,m_{s_{n-1}})$ of $n$ integers that we call {\bf offsets}, with $n$ the number of routes in $\cal R_{\cal C}$. The number $m_{s_i}$ represents the index of the first slot used in a period  by the route $r_{s_i} \in \cal R_{\cal C}$ at its source $s_i$.
      A $(P,\tau)$-periodic affectation must have no {\bf collision} between two routes in $\cal R_{\cal C}$, that is $\forall (r_{s_i}, r_{s_j}) \in {\cal R_{\cal C}}^2, i \neq j$, % with $\tau$ the size (in number of consecutive slots) of each message that must be periodically sent on each route of ${\cal R}_{\cal C}$, 
      we have $$[t(u,r_{s_i})] \cap [t(u,r_{s_j})] = \emptyset .$$

%       Notice that the notion of $P$-periodic affectation \textbf{is not monotone} with regard to $P$. 
      As an example of a $(2,1)$-periodic affectation, let consider a routage graph with routes $\{r_{s_i}\}_{i=0,\dots,n-1}$, such that all pairs of routes intersect at a different edge.
      We set $\tau = 1$ and the delays are chosen so that if $r_{s_i}$ and $r_{s_j}$ have $v$ as first common vertex then we have $\lambda(v,r_{s_i}) - \lambda(v,r_{s_j})=1$.There is a $(2,1)$-periodic affectation by setting all $m_{s_i}$ to $0$.

%       A {\bf conflict graph} represents the collision between the routes of $ {\cal R}_{\cal C}$. The vertices of a conflict graph $G = (V,E)$ are the routes of $ {\cal R}_{\cal C}$, and there is an edge between two vertices if and only if there is a common arc between the two routes in $ {\cal R}_{\cal C}$.
%       
%       Given $u$ and $v$ two vertices of the conflict graph, corresponding to two routes colliding in $ {\cal R}_{\cal C}$. The weight of an edge, $w(u,v)$, is the absolute value of the difference between the distance of the two routes between their respective source node and the collision point.
%       
%       A labeling $F$ of such a graph is an affectation of an integer to each vertex, such that for each vertex $u$, $f(u) \neq f(v)+w(u,v)\mod P$, where $v$ are the neighbors of $u$ in the conflict graph and $P$ our period.

      \begin{figure}[ht]
       \begin{center}

@@

\tikzset{
  LabelStyle/.style = { rectangle, rounded corners, draw,
                       font = \bfseries },
  EdgeStyle/.append style = {->} }
  \SetGraphUnit{5}
  \node[draw,circle] (s3) at (4, 2) {$s_2$}; 
  \node[draw,circle] (s2) at (0, 4) {$s_1$}; 
  \node[draw,circle] (s1) at (0, 6) {$s_0$};

  \node[draw,circle] (t3) at (14, 7) {$t_2$}; 
  \node[draw,circle] (t2) at (14, 4) {$t_1$}; 
  \node[draw,circle] (t1) at (10, 2) {$t_0$};

@@

\end{tikzpicture}
      \end{center}
       \caption{A routage graph with $(0,0,0)$ as a $(2,1)$-periodic affectation}
      \end{figure}

@@

    We want to ensure that there is an affectation which allows to send periodic messages from sources to target
    without collisions. The problem we need to solve is thus the following:

@@

      \noindent {\bf Input:} a routage graph $\cal R_{\cal C}$, an integer $\tau$ and an integer $P$.

      \noindent {\bf Question:} does there exist a $(P,\tau)$-periodic affectation of $\cal R_{\cal C}$ ?

      We will prove in Sec.~\ref{sec:complexity} that the problem PRA is $\NP$-complete, even in restricted settings.
      Even approximating the smallest value of $P$ for which there is a $(P,\tau)$-periodic affectation is hard.

      An unusual property of affectation is that given a routage graph, we may have a $(P,\tau)$-periodic affectation but no
      $(P',\tau)$-periodic affectation with $P' > P$: the existence of an affectation is not monotone with regards to $P$.

	\begin{lemma} \label{lemma:monotonic}
	 For any odd $P$, there is a routage graph such that there is $(2,1)$-periodic affectation but no $(P,1)$-periodic affectation.
	\end{lemma}
\begin{proof}

      Consider the routage graph ${\cal R_{\cal C}}$ given in the previous subsection. 
      We change the delays so that for $v$, the first vertex which belongs to $r_i$ and $r_j$,
      we have $\lambda(v,r_i) - \lambda(v,r_j)= P$, where $P$ is an odd number smaller than $n$, the number of routes in ${\cal R_{\cal C}}$. In such a graph, there is no $(P,\tau)$-periodic affectation, since the problem reduces to finding a $P$-coloring in a complete graph with $n > P$ vertices.\\
      If we consider a period of $2$, for all $i \neq j$, $\lambda(v,r_i) - \lambda(v,r_j) \mod 2 = 1$ . Therefore $(0,\dots,0)$ is a $(2,1)$-periodic affectation of ${\cal R_{\cal C}}$.

@@

      In the context of cloud-RAN applications, we consider here the digraph $G=(V,A)$ modeling the target network 
      and two disjoint subsets of vertices $S$ and $T$ of equal cardinality, where $S$ is the set of RRHs and $T$ is the set of BBUs. 
      A \textbf{symmetric} assignment ${\cal C}$, is an involutive function from $S$ to $T$, which maps each element $s\in S$ to ${\cal C}(s)$ and ${\cal C}(s)$ to $s$. It can also be seen as the set of pairs $(s,{\cal C}(s))$ and $({\cal C}(s),s)$.\\
      The routing function ${\cal R}$ associates to each couple $(s,{\cal C}(s))$ a route called $r_s$ and to each couples $({\cal C}(s),s)$ a route called $r_{{\cal C}(s)}$.\\     
       We are given a period $P$, a routing function ${\cal R}$, and we consider a $(P,\tau)$-periodic affectation of ${\cal R}_{\cal C}$ which associates $m_s$ to the route $r_s$ which begins by $s$, and $m_{{\cal C}(s)}$ to the routes $r_{{\cal C}(s)}$ which begins by ${\cal C}(s)$.  This affectation represents the following process: first a message is sent at $s$, through the route $r_s$, at time $m_s$.

%       
%       We denote by $n$ the size of $S$ and $L$. We are given a period $P$, a routing function ${\cal R}$ and a bijection $\rho:L\rightarrow S$ which assigns a BBU to each RRH. Let ${\cal C}_{\rho} = \{(l,\rho(l))\}_{l \in L} \cup \{(s,\rho^{-1}(s))\}_{s \in S}$. Let consider a $P$-periodic affectation of ${\cal C}_{\rho}$ which associates $m_l$ to 
%       $(l,\rho(l))$ and $m_{\rho(l)}$ to $(\rho(l),l)$.

@@

      This message is received by ${\cal C}(s)$ at time $t({\cal C}(s),r_s)$. It is then sent back to $s$ in the same period at time $m_{{\cal C}(s)}$ if $m_{{\cal C}(s)} > t({\cal C}(s),r_s)$, otherwise at time $m_{{\cal C}(s)}$ in the next period. The time between the arrival of the message and the time it is sent back is called the \textbf{waiting time} and is defined by $w_s = m_{{\cal C}(s)} - t({\cal C}(s),r_s)$ if $m_{{\cal C}(s)} > t({\cal C}(s),r_s)$ and $w_s = m_{{\cal C}(s)} + P - t({\cal C}(s),r_s)$ otherwise.

@@

      Thus, the whole process time for a message sent at vertex $s$ is equal to
      $$
      PT(s)=\lambda(r_s)+ w_s+\lambda(r_{{\cal C}(s)}).
      $$

@@

    The {\bf maximum process time} of the $(P,\tau)$-periodic affectation ${\cal M} $ is defined by $MT({\cal M})=\max\limits_{s \in S} PT(s)$. The problem we want to solve is the following.

      \noindent {\bf Periodic Assignment for Low Latency(PALL)}

      \noindent {\bf Input:}  A routage graph ${\cal R}_{\cal C}$ with ${\cal C}$ a symmetric assignment, a period $P$, an integer $\tau$, an integer $T_{max}$.

      \noindent {\bf Question:} does there exist a $(P,\tau)$-periodic affectation ${\cal M}$ of ${\cal R}_{\cal C}$ such that $MT({\cal M}) \leq T_{max}$?

      \todo{Si on veut, on peut parler du temps moyen aussi ici, seulement si on fait quelque chose dessus dans la suite}
      %The related optimisation problem we will focus on  consists in minimizing  $MT({\cal M})$. Note that in the context of cloud-RAN networks, we consider $P=1ms$, $\theta=2.6ms$ and $T_{max}$ must be less or equal to $3ms$.
      %cette remarque doit être dans la partie expérimentale

  In this article, we works on a given routage, but we could imagine a problem, in which, considering a given graph $G = (V,A)$ and a period P, the question is to find, if it exists, a routage graph ${\cal R}_{\cal C}$, in which there is $(P,\tau)$-periodic affectation.

  \todo{Dire si on regle le probleme de trouver un bon routage dans un cas particulier après}

@@

 In this section we assume that the size of a message $\tau$ is equal to one. 
 We will prove the hardness of PRA and PALL for this parameter, which implies the hardness of the general problems. 
Consider an instance of the problem PRA, i.e., a routage graph $\cal{R_{\cal C}}$ and a period $P$. \\
The {\bf conflict depth} of a route is the number of other routes which share an edge with it. 
The conflict depth of a routage graph  $\cal{R_{\cal C}}$ is the maximum of the conflict depth of the routes in $\cal{R_{\cal C}}$.\\
The {\bf load} of a routage graph is the maximal number of routes sharing the same arc.
It is clear that a $(P,1)$-periodic affectation must satisfy that $P$ is larger or equal to the load.

@@

 \begin{proposition}
Problem PRA is $\NP$-complete, when the routing is of conflict depth two.
\end{proposition}
 \begin{proof}
 The problem $PRA$ is in $\NP$ since given an offset for each route in an affectation, it is easy to check in linear time in the number of edges whether there are collisions.

  Let $H=(V,E)$ be a graph and let $d$ be its maximal degree. We consider the problem to determine whether $H$ is edge-colorable
  with $d$ or $d+1$ colors. The edge coloring problem is $\NP$-hard~\cite{holyer1981np} and we reduce it to PRA to prove its $\NP$-hardness. We define from $H$ an instance of PRA as follows. 
  For each $v$ in $V$, the graph $G$ has two vertices $v_1, v_2$, and for each $(u,v) \in E$,the graph G has two vertices $s_{u,v}, t_{u,v}$. Set an arbitrary orientation of the edge $(u,v)$, such that the following route is directed from $u$ to $v$.

%   The set $A$ of arcs of $G$ is: 
%   $$ \{(v_1,v_2) \mid v\in V\} \cup \{(u_2,v_1)\mid u \neq v \in V^2\} \cup \{(s_{u,v},u_1),(v_2,t_{u,v}) \mid (u,v) \in E \}. $$
  For each edge $(u,v) \in E$, there is a route $s_{u,v},u_1,u_2,v_1,v_2,t_{u,v}$ in ${\cal R}$.  
  All these arcs are of weight $0$. 
  The set of arcs of G is the union between all the arcs of the previous routes.
  The affectation ${\cal C}$ is the set of pair of vertices $(s_{u,v}, t_{u,v})$.

  Observe that the existence of a $d$-coloring of $H$ is equivalent to the existence of a $(d,1)$-periodic affectation
  of ${\cal R}_{\cal C}$. Indeed, a $d$-coloring of $H$ can be seen as a labeling of its edges by the integers
  in $\{0,\dots,d-1\}$ and we have a bijection between $d$-colorings of $H$ and offsets of the routes of ${\cal R}_{\cal C}$.
  By construction, the constraint of having no collision between the routes is equivalent to the fact that no two adjacent edges have
  the same color. Therefore we have reduced edge coloring to PRA which concludes the proof. 
 \end{proof}
 \todo{Faire un dessin d'illustration ?}

@@

\begin{corollary}
Problem PALL is $\NP$-complete for routing of conflict depth two.
\end{corollary}
\begin{proof}
 We consider $({\cal R}_{\cal C},P,\tau)$ an instance of $PRA$. We assume that no vertex appears both in the first and second position in a pair of ${\cal C}$. Remark that this condition is satisfied in the previous proof, which makes the problem $PRA$ restricted to this kind of instances $\NP$-complete. 
 Let us define $T_{max} = 2 \times \max_{r \in {\cal R}} \lambda(r) + P$. We consider ${\cal C}'$ and ${\cal R}'_{\cal C'}$ symmetrized versions of ${\cal C}$ and  ${\cal R}_{\cal C}$ where for every route there is a symmetric route with new arcs and the same weights.
 The instance $({\cal R'}_{\cal C'},P,\tau,T_{max})$ is in PALL if and only if $({\cal R}_{\cal C},P,\tau)	$
 is in $PRA$. Indeed the waiting time of each route is by definition less than $P$ and thus the maximal process time is always less than $T_{max}$. Moreover a $(P,\tau)$-affectation of ${\cal R}_{\cal C}$ can be extended into a $(P,\tau)$-affectation of ${\cal R'}_{\cal C'}$ in the following way. For each route $r_{u,v}$, the time at which the message arrives is $t(v,r_{u,v})$, then we choose as offset for $r_{v,u}$, $-t(v,r_{u,v}) \mod P$. The symmetry ensures that each new route $r_{v,u}$ in ${\cal R'}_{\cal C'}$ uses the same times slot as $r_{u,v}$ and thus avoid collisions.
\end{proof}

Let MIN-PRA be the problem, given a routage graph and an assignment, to find the minimal period $P$ such that there is a $P$-periodic affectation.

>>>>

%opening
\title{How to Schedule Periodic Messages in a Network Without Congestion}
\author{DB,CC,MG,OM,YS}

@@

\begin{abstract}
We study the sending of periodic messages in a graph following predefined paths, a problem arising from the communication between radio antennas and their signal processing units located in distant data-centers. The messages must be sent deterministically to avoid collisions on the network and thus to minimize the latency. We prove that the problem is $\NP$-complete even for very restricted graphs. We then deal with the case of a star graph and for different regimes of the parameters we propose algorithms that are experimentally shown to work extremely well
compared to traditional statistical multiplexing.
\end{abstract}

\section{Introduction}

@@

 The network topology is modeled by a directed weighted graph and a set of paths (routes) from source nodes (RRHs) to target nodes (BBUs). The time is discretized and a unit of time corresponds to the transmission of a minimal unit of data over the network.
 The process must be periodic, that is if we look at the network at time $t$ and $t+P$ where $P$ is the period, all messages are at the same position in the network. Moreover, we want to avoid any  buffering in internal nodes of the graph. Hence we need to design 
 a periodic process with no collision between messages. To ensure that this property is always true we propose a deterministic process.  We have two sets of parameters that we can choose when building a periodic sending process, called a \textbf{periodic assignment} in this article: the time at which each packet is sent by a RRH in each period and the waiting time in the BBU before the answer is sent back to the RRH. When building a periodic assignment, we must take into account the periodicity which makes many scheduling methods inapplicable. Not only a message must not collide with the other messages sent by the others BBU/RRH in the same period, but also in the previous and following periods. Moreover the latency must be minimized, that is the time 
 between the emission of a message and the complete return of its answer. It means that the only buffering we are allowed -- the waiting time before sending back the answer-- must be small, in particular when the route is long.

@@

 In Sec.~\ref{sec:def} we propose a model of the network and the periodic sending of messages along its routes. 
 We then formalize our two problems of finding a periodic assignment for sending messages without collisions: PRA (sending of the message only) and PALL (sending of the message and answer).  
In Sec.~\ref{sec:complexity}, we prove that the problem PRA and PALL are $\NP$-hard and cannot be approximated even for very
restricted graphs. 
Therefore in the next two sections, we study a simple but very common topology, that we call the
star topology, on which the messages can collide on only one arc.
In Sec.\ref{sec:PAZL}, we study a variant of PALL called PAZL  where the waiting times must all be zero. We provide several algorithms, some of their theoretical properties and give experimental evidences that they find periodic assignments when the network is no too loaded. Finally in Sec.\ref{sec:PALL}, we propose algorithms for the general PALL problem and exhibit one which works extremely well in our experiments even in loaded networks. In particular we show that the deterministic communication schemes we design vastly outperform the traditional stochastic multiplexing with regards to latency.

\section{Model and problems}\label{sec:def}

  \subsection{Network modeling}

The network is modeled as a directed graph $G=(V,A)$. Each arc  $(u,v)$ in $A$ is labeled by an integer $dl(u,v) \geq 1$ that we call the delay and which represents the time slots taken by a signal to go from $u$ to $v$ using this arc. A {\bf route} $r$ in $G$ is a path, that is sequence of consecutive arcs $a_0, \ldots , a_{k-1}$, with $a_i=(u_i,u_{i+1}) \in A$.  The {\bf latency} of a vertex $u_i$ in a path $r$, with $i \geq 1$, is defined by $\lambda(u_i,r)= \sum\limits_{0 \leq j <i} dl(a_j)$. We also define $\lambda(u_0,r)=0$. The latency of the route $r$ is defined by $\lambda (r)= \lambda (u_k,r)$.
We denote by $\cal R$ a set of routes in $G$ such that no two routes have the same first vertices or the same last vertices.
The pair $(G,\cal R)$ is called a {\bf routage graph} and represents our network.
The first vertex of a route models an antenna (RRH) and the last one a data-center (BBU) which processes the messages sent by the antenna. We denote by $\rho$ the bijection induced by ${\cal R}$, $\rho(s) = t$ if $s$ is the first vertex of a route $r$ and $t$ the last.

   \subsection{Slotted time Model}

      Consider a route $r$ from $s$ to $t$. If we send a message at some point in time $m$ from $s$,
      then it will arrive at the vertex $v$ in $r$ at time $m + \lambda(v,r)$. Consider now a positive integer $P$ called the {\bf period}. In our problem, we send messages in the network with period $P$. The time will thus be cut into slices of $P$ discrete slots and we need to consider only one of such slices since they will all be identical. 
       Hence we define the first time slot at which a message sent from $s$ will reach a vertex $v$ in $r$ by $t(v,r) = m + \lambda(v,r) \mod P$.

      A message usually cannot be transported in a single time slot. We denote by $\tau$ the number 
      of \emph{consecutive slots} necessary to transmit a message. Let us call $[t(v,r)]_{P,\tau}$ the set of time slots used by $r$ at a vertex $v$ in a period $P$, that is $[t(v,r)]_{P,\tau} = \{t(v,r) + k \mod P \mid 0 \leq k < \tau \}$. Usually $P$ and $\tau$ will be clear from the context and we will denote $[t(v,r)]_{P,\tau}$ by $[t(v,r)]$.

      A {\bf $(P,\tau)$-periodic assignment} of a routage graph $(G,\cal R)$ is a sequence  ${\cal M}=(m_1, \ldots ,m_n)$ of $n$ integers that we call {\bf offsets}, with $n$ the number of routes in $\cal R$. The number $m_{i}$ represents the time at which a message is emitted at the first vertex of the route $r_{i}$. We will always consider that the $m_{i}$ are between $O$ and $P-1$. A $(P,\tau)$-periodic assignment must have no {\bf collision} between two routes in $\cal R$, that is $\forall (r_{i}, r_{j}) \in {\cal R}^2, i \neq j$, % with $\tau$ the size (in number of consecutive slots) of each message that must be periodically sent on each route of ${\cal R}_{\cal C}$, 
      we have $$[t(u,r_{i})] \cap [t(u,r_{j})] = \emptyset .$$

%       Notice that the notion of $P$-periodic assignment \textbf{is not monotone} with regard to $P$. 
      As an example of a $(2,1)$-periodic assignment, let consider a routage graph with routes $\{r_{i}\}_{i=1,\dots,n}$, such that all pairs of routes intersect at a different edge.
      We set $\tau = 1$ and the delays are chosen so that if $r_{i}$ and $r_{j}$ have $v$ as first common vertex then we have $\lambda(v,r_{i}) - \lambda(v,r_{j})=1$.There is a $(2,1)$-periodic assignment by setting all $m_{i}$ to $0$.

%       A {\bf conflict graph} represents the collision between the routes of $ {\cal R}_{\cal C}$. The vertices of a conflict graph $G = (V,E)$ are the routes of $ {\cal R}_{\cal C}$, and there is an edge between two vertices if and only if there is a common arc between the two routes in $ {\cal R}_{\cal C}$.
%       
%       Given $u$ and $v$ two vertices of the conflict graph, corresponding to two routes colliding in $ {\cal R}_{\cal C}$. The weight of an edge, $w(u,v)$, is the absolute value of the difference between the distance of the two routes between their respective source node and the collision point.
%       
%       A labeling $F$ of such a graph is an assignment of an integer to each vertex, such that for each vertex $u$, $f(u) \neq f(v)+w(u,v)\mod P$, where $v$ are the neighbors of $u$ in the conflict graph and $P$ our period.

      \begin{figure}[ht]
      % \begin{center}

@@

\tikzset{
  LabelStyle/.style = { rectangle, rounded corners, draw,
                       font = \bfseries },
  EdgeStyle/.append style = {->} }
  \SetGraphUnit{5}
  \node[draw,circle] (s3) at (4, 2) {$s_3$}; 
  \node[draw,circle] (s2) at (0, 4) {$s_2$}; 
  \node[draw,circle] (s1) at (0, 6) {$s_1$};

  \node[draw,circle] (t3) at (14, 7) {$t_3$}; 
  \node[draw,circle] (t2) at (14, 4) {$t_2$}; 
  \node[draw,circle] (t1) at (10, 2) {$t_1$};

@@

\end{tikzpicture}
     % \end{center}
       \caption{A routage graph with $(0,0,0)$ as a $(2,1)$-periodic assignment}
      \end{figure}

@@

    We want to find an assignment which allows to send periodic messages from sources to targets
    without collisions. We introduce the following associated decision problem, useful for hardness proofs.

@@

      \noindent {\bf Input:} a routage graph $(G,\cal R)$, an integer $\tau$ and an integer $P$.

      \noindent {\bf Question:} does there exist a $(P,\tau)$-periodic assignment of $(G,\cal R)$ ?

      We will prove in Sec.~\ref{sec:complexity} that the problem PRA is $\NP$-complete, even in restricted settings.
      Even approximating the smallest value of $P$ for which there is a $(P,\tau)$-periodic assignment is hard.

      An unusual property of assignment is that given a routage graph, we may have a $(P,\tau)$-periodic assignment but no
      $(P',\tau)$-periodic assignment with $P' > P$: the existence of an assignment is not monotone with regards to $P$.

	\begin{lemma} \label{lemma:monotonic}
	 For any odd $P$, there is a routage graph such that there is $(2,1)$-periodic assignment but no $(P,1)$-periodic assignment.
	\end{lemma}
\begin{proof}

      Consider the routage graph $(G,{\cal R})$ given in the previous subsection. 
      We change the delays so that for $v$, the first vertex which belongs to $r_i$ and $r_j$,
      we have $\lambda(v,r_i) - \lambda(v,r_j)= P$, where $P$ is an odd number smaller than $n$, the number of routes in ${\cal R_{\cal C}}$. In such a graph, there is no $(P,\tau)$-periodic assignment, since the problem reduces to finding a $P$-coloring in a complete graph with $n > P$ vertices.\\
      If we consider a period of $2$, for all $i \neq j$, $\lambda(v,r_i) - \lambda(v,r_j) \mod 2 = 1$ . Therefore $(0,\dots,0)$ is a $(2,1)$-periodic assignment of ${\cal R}$.

@@

      In the context of cloud-RAN applications, we need to send message from a RRH $s$ to a BBUs $\rho(s)$ and then 
      we must send the answer from $\rho(s)$ back to $s$. We say that a routage graph $(G, {\cal R})$ is \textbf{symmetric}
      if for all route $r$ with first vertex $s$ and last vertex $target(s)$ there is a route $r'$ with first vertex $target(s)$
      and last vertex $s$. In all practical cases the routes $r$ and $r'$ will be the same with the orientation of the arcs reversed, which corresponds to bidirectional links in the networks, but we will not enforce this property.
      If the first vertex of $r$ is $s$, we will denote by $\rho(r)$ the route whose first vertex is $\rho(s)$.

         We now give a new interpretation of a $(P,\tau)$-periodic assignment of a $(G,{\cal R})$ symmetric routing graph, so that it represents the sending of a message and of its answer.
	This assignment represents the following process: first a message is sent at $s$, through the route $r$, at time $m$.

%       
%       We denote by $n$ the size of $S$ and $L$. We are given a period $P$, a routing function ${\cal R}$ and a bijection $\rho:L\rightarrow S$ which assigns a BBU to each RRH. Let ${\cal C}_{\rho} = \{(l,\rho(l))\}_{l \in L} \cup \{(s,\rho^{-1}(s))\}_{s \in S}$. Let consider a $P$-periodic assignment of ${\cal C}_{\rho}$ which associates $m_l$ to 
%       $(l,\rho(l))$ and $m_{\rho(l)}$ to $(\rho(l),l)$.

@@

      This message is received by $\rho(s)$ at time $t(\rho(s),r)$. It is then sent back to $s$ on the route $\rho(r)$ in the same period at time $m_{\rho(s)}$ if $m_{\rho(s)} > t(\rho(s),r)$, otherwise at time $m_{\rho(s)}$ in the next period. The time between the arrival of the message and the time it is sent back is called the \textbf{waiting time} and is defined by $w_s = m_{\rho(s)} - t(\rho(s),r)$ if $m_{\rho(s)} > t(\rho(s),r)$ and $w_s = m_{\rho(s)} + P - t(\rho(s),r)$ otherwise.

@@

      Thus, the whole process time for a message sent at vertex $s$ on the route $r$ is equal to
      $$
      PT(s)=\lambda(r)+ w_s+\lambda(r).
      $$

@@

    The {\bf maximum process time} of the $(P,\tau)$-periodic assignment ${\cal M} $ is defined by $MT({\cal M})=\max\limits_{s \in S} PT(s)$. We consider the following decision problem.

      \noindent {\bf Periodic Assignment for Low Latency (PALL)}

      \noindent {\bf Input:}  A symmetric routage graph $(G,{\cal R})$, a period $P$, an integer $\tau$, an integer $T_{max}$.

      \noindent {\bf Question:} does there exist a $(P,\tau)$-periodic assignment ${\cal M}$ of $(G,{\cal R})$ such that $MT({\cal M}) \leq T_{max}$?

      As a consequence of the $\NP$-hardness of PRA, we show in the next subsection that this problem is $\NP$-hard. 
      In Sec.~\ref{sec:PALL} we will study heuristics which try to solve the search version of PALL (computing an assignment), also denoted by PALL for simplicity.

@@

 In this section we always assume that the size of a message $\tau$ is equal to one. 
 We will prove the hardness of PRA and PALL for $\tau =1$ which implies the hardness of the general problems. 
Consider an instance of the problem PRA, i.e., a routage graph $(G,\cal{R})$ and a period $P$.
The {\bf conflict depth} of a route is the number of other routes which share an edge with it. 
The conflict depth of a routage graph $(G,\cal{R})$ is the maximum of the conflict depth of the routes in $\cal{R}$.
The {\bf load} of a routage graph is the maximal number of routes sharing the same arc.
Remark that a $(P,1)$-periodic assignment must satisfy that $P$ is larger or equal to the load.

@@

 \begin{proposition}
Problem PRA is $\NP$-complete, when the routing is of conflict depth two.
\end{proposition}
 \begin{proof}
 The problem $PRA$ is in $\NP$ since given an offset for each route in an assignment, it is easy to check in linear time in the number of edges whether there are collisions.

  Let $H=(V,E)$ be a graph and let $d$ be its maximal degree. We consider the problem to determine whether $H$ is edge-colorable
  with $d$ or $d+1$ colors. The edge coloring problem is $\NP$-hard~\cite{holyer1981np} and we reduce it to PRA to prove its $\NP$-hardness. We define from $H$ an instance of PRA as follows. 
  For each $v$ in $V$, the graph $G$ has two vertices $v_1, v_2$, and for each $(u,v) \in E$,the graph G has two vertices $s_{u,v}, t_{u,v}$. Set an arbitrary orientation of the edge $(u,v)$, such that the following route is directed from $u$ to $v$.
%   The set $A$ of arcs of $G$ is: 
%   $$ \{(v_1,v_2) \mid v\in V\} \cup \{(u_2,v_1)\mid u \neq v \in V^2\} \cup \{(s_{u,v},u_1),(v_2,t_{u,v}) \mid (u,v) \in E \}. $$
  For each edge $(u,v) \in E$, there is a route $s_{u,v},u_1,u_2,v_1,v_2,t_{u,v}$ in ${\cal R}$.  
  All these arcs are of weight $0$. The set of arcs of G is the union between all the arcs of the previous routes.

  Observe that the existence of a $d$-coloring of $H$ is equivalent to the existence of a $(d,1)$-periodic assignment
  of $(G,{\cal R})$. Indeed, a $d$-coloring of $H$ can be seen as a labeling of its edges by the integers
  in $\{0,\dots,d-1\}$ and we have a bijection between $d$-colorings of $H$ and offsets of the routes of ${\cal R}$.
  By construction, the constraint of having no collision between the routes is equivalent to the fact that no two adjacent edges have the same color. Therefore we have reduced edge coloring to PRA which concludes the proof. 
 \end{proof}

@@

\begin{corollary}
Problem PALL is $\NP$-complete for routing of conflict depth two.
\end{corollary}
\begin{proof}
 We consider $((G,{\cal R}),P,\tau)$ an instance of $PRA$. We assume that no vertex is the first of some route and the last of another one. Remark that this condition is satisfied in the previous proof, which makes the problem $PRA$ restricted to this kind of instance $\NP$-complete. 
 Let us define $T_{max} = 2 \times \max_{r \in {\cal R}} \lambda(r) + P$. We define $(G',{\cal R}')$ a symmetric routage graph from $(G,{\cal R})$ where for every route we add a symmetric route with new arcs of opposite orientation and the same weights.
 The instance $((G',{\cal R'}),P,\tau,T_{max})$ is in PALL if and only if $((G,{\cal R}),P,\tau)$ is in $PRA$. Indeed the waiting time of each route is by definition less than $P$ and thus the maximal process time is always less than $T_{max}$. Moreover a $(P,\tau)$-assignment of $(G,{\cal R})$ can be extended into a $(P,\tau)$-assignment of $(G',{\cal R'})$ in the following way. For each route $r \in \cal{R}$ starting at $s$, the time at which the message arrives is $t(\rho(s),r)$, then we choose as offset for $\rho(r) \in \cal{R}'$ $-t(\rho(s),r) \mod P$. The symmetry ensures that each new route $\rho(r)$ in ${\cal R'}$ uses the same times slot as $r$ one each of its node and thus avoid collisions.
\end{proof}

Let MIN-PRA be the problem, given a routage graph and an assignment, to find the minimal period $P$ such that there is a $P$-periodic assignment.
